---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Download into 'dataset.zip / extract into 'activity.csv'
```{r, echo=TRUE}
fileurl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileurl, destfile = "dataset.zip", method = "curl")
unzip("dataset.zip")
activity <- read.csv("activity.csv")
```

Add a column to specify the day of the week
```{r, echo=TRUE}
activity$day <- weekdays(as.Date(activity$date))
```


## What is mean total number of steps taken per day?
Get the total amount of steps for each day:
```{r, echo=TRUE}
totalStepsByDay <- aggregate(activity$steps, by=list(Category=activity$date), FUN=sum)
```
Histogram:
```{r, echo=TRUE}
qplot(date, steps, data=activity, stat="summary", fun.y="sum", geom="bar")
```

Get the mean an median of steps by day:
```{r, echo=TRUE}
summary(totalStepsByDay$x, na.rm=TRUE)
Min.  1st Qu.  Median   Mean   3rd Qu.  Max.    NA's 
41    8841     10760    10770  13290    21190   8 
```
The **mean** total number of steps taken per day is **10770**
The **median** total number of steps taken per day is **10760**


## What is the average daily activity pattern?
Get the mean amount of steps for each interval:
```{r, echo=TRUE}
totalStepsByInterval <- aggregate(activity$steps, by=list(Category=activity$interval), FUN=mean, na.rm=TRUE)
```
Get the interval with the most steps
```{r, echo=TRUE}
totalStepsByInterval[which.max(totalStepsByInterval$x),]
```
The 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps is:
**835** with **206.1698** steps on average


## Imputing missing values
Total number of missing values in the dataset: **2304**
```{r, echo=TRUE}
sum(is.na(activity$steps))
```
New dataset, filling gaps using mean:
```{r, echo=TRUE}
noNAactivity <- activity
for (i in 1:17568) {
  if (is.na(noNAactivity$steps[i])== TRUE) {
    noNAactivity$steps[i] <- round(meanStepsByInterval[meanStepsByInterval[,1] == noNAactivity$interval[i],2])
  }
}
```
Old dataset: 
head(activity)
  steps       date interval    day
1    NA 2012-10-01        0 Monday
2    NA 2012-10-01        5 Monday
3    NA 2012-10-01       10 Monday
4    NA 2012-10-01       15 Monday
5    NA 2012-10-01       20 Monday
6    NA 2012-10-01       25 Monday
summary(totalStepsByDay$x)
Min.  1st Qu.  Median   Mean   3rd Qu.  Max.    NA's 
41    8841     10760    10770  13290    21190   8 

New dataset:
head(noNAactivity)
  steps       date interval    day
1     2 2012-10-01        0 Monday
2     0 2012-10-01        5 Monday
3     0 2012-10-01       10 Monday
4     0 2012-10-01       15 Monday
5     0 2012-10-01       20 Monday
6     2 2012-10-01       25 Monday
summary(totalStepsByDayFILLED$x)
Min.  1st Qu.  Median   Mean   3rd Qu.   Max.    NA's 
41    8841     10760    10770  13290     21190   8 

Histogram:
```{r, echo=TRUE}
qplot(date, steps, data=noNAactivity, stat="summary", fun.y="sum", geom="bar")
```

## Are there differences in activity patterns between weekdays and weekends?
